### 2020-08-09 MSc-dissertation.tex
New structure

\documentclass[12pt,a4paper]{report}

\usepackage{epcc}
\usepackage{graphics}
\usepackage{hyperref}
% \usepackage{mathtools}
\usepackage{amsmath}
\usepackage{float}

\begin{document}

\pagenumbering{roman}

\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}

\author{Chaolin Han}

\date{\today}

\makeEPCCtitle

\thispagestyle{empty}

\vspace{11cm}

\begin{center}

    \large{MSc in High Performance Computing}

    \large{The University of Edinburgh}

    \large{Year of Presentation: 2020}

\end{center}

\newpage

\begin{abstract}
    This is the bit where you summarise what is in your thesis.
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listoftables
\listoffigures

% \begin{titlepage}
%     \vspace*{2in}
%     % an acknowledgements section is completely optional but if you decide
%     % not to include it you should still include an empty {titlepage}
%     % environment as this initialises things like section and page numbering.
%     \section*{Acknowledgements}

%     This template is a slightly modified version of the one developed by
%     Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
%     acknowledgement follows:

%     {\em This template has been produced with help from many former students who
%     have shown different ways of doing things. Please make suggestions for
%     further improvements.}

% \end{titlepage}

\pagenumbering{arabic}

\input{introduction/introduction.tex}

\input{background/background.tex}

\input{modelling/modelling.tex}

\input{implementation/implementation.tex}

\input{performance/performance.tex}

\input{conclusion/future_works.tex}

\input{conclusion/conclusion.tex}

% \chapter{Results and Discussions}

% [WHAT results is outputted and WHAT topics are discussed]

% \section{ABC SMC results}

% [inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]

% [model selections result; bayes factor]

% \section{Discussions}

% [goodness of fit (evaluation); preferred models and effects of modification to basic model]

% [efficiency and trade-offs in scaling-up]

% [POSSIBLE: compare to exact inference; compared to other implementations]

% [limitations]

\input{appendix/appendix.tex}

\begin{thebibliography}{100}

    \bibitem{axonal} Becker, T., Wullimann, M.F., Becker, C.G., Bernhardt, R.R. and Schachner, M., 1997. Axonal regrowth after spinal cord transection in adult zebrafish. Journal of Comparative Neurology, 377(4), pp.577-595.

    \bibitem{ref:Tsarouchas} Tsarouchas, T.M., Wehner, D., Cavone, L., Munir, T., Keatinge, M., Lambertus, M., Underhill, A., Barrett, T., Kassapis, E., Ogryzko, N. and Feng, Y., 2018. Dynamic control of proinflammatory cytokines Il-1$\beta$ and Tnf-$\alpha$ by macrophages in zebrafish spinal cord regeneration. Nature communications, 9(1), pp.1-17.

    \bibitem{Anderson} Anderson, W.D., Makadia, H.K., Greenhalgh, A.D., Schwaber, J.S., David, S. and Vadigepalli, R., 2015. Computational modeling of cytokine signaling in microglia. Molecular BioSystems, 11(12), pp.3332-3346.

    \bibitem{ref:pyabc} Klinger, E., Rickert, D. and Hasenauer, J., 2018. pyABC: distributed, likelihood-free inference. Bioinformatics, 34(20), pp.3591-3593.

    \bibitem{ref:kernel} Filippi, S., Barnes, C. P., Cornebise, J., and Stumpf, M..H. (2013). On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo. Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.

    \bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014. A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation. Nature Protocols. 9 (2) pp. 439–456.

    \bibitem{ref:MCMC} Gilks, W.R., Richardson, S. and Spiegelhalter, D.J., 1996. Markov Chain Monte Carlo in Practice Chapman \& Hall: London. Markov Chain Monte Carlo in practice. Chapman and Hall, London.

    \bibitem{ABC_rejection} Tavaré, S., Balding, D.J., Griffiths, R.C. and Donnelly, P., 1997. Inferring coalescence times from DNA sequence data. Genetics, 145(2), pp.505-518.

    \bibitem{summaryD} Joyce, P. and Marjoram, P., 2008. Approximately sufficient statistics and Bayesian computation. Statistical applications in genetics and molecular biology, 7(1).

    \bibitem{summaryD2} Nunes, M.A. and Balding, D.J., 2010. On optimal selection of summary statistics for approximate Bayesian computation. Statistical applications in genetics and molecular biology, 9(1).

    \bibitem{threshold} Silk, D., Filippi, S. and Stumpf, M.P., 2012. Optimizing threshold-schedules for approximate Bayesian computation sequential Monte Carlo samplers: applications to molecular systems. arXiv preprint arXiv:1210.3296.

    \bibitem{population} Klinger, E. and Hasenauer, J., 2017, September. A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo. In International Conference on Computational Methods in Systems Biology (pp. 128-144). Springer, Cham.

    \bibitem{model_compare} Toni, T. and Stumpf, M.P., 2010. Simulation-based model selection for dynamical systems in systems and population biology. Bioinformatics, 26(1), pp.104-110.

    \bibitem{cosmology} Weyant, A., Schafer, C. and Wood-Vasey, W.M., 2013. Likelihood-free cosmological inference with type Ia supernovae: approximate Bayesian computation for a complete treatment of uncertainty. The Astrophysical Journal, 764(2), p.116.

    \bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C. 2017. Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J. R. Soc. Interface14: 20170340.

    \bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.

    \bibitem{ref:adpt_dis} Prangle, D., 2017. Adapting the ABC distance function. Bayesian Analysis, 12(1), pp.289-309.

    \bibitem{ref:adpt_pop} Klinger, E. and Hasenauer, J., 2017, September. A scheme for adaptive selection of population sizes in approximate Bayesian computation-sequential Monte Carlo. In International Conference on Computational Methods in Systems Biology (pp. 128-144). Springer, Cham.

    \bibitem{Toni} Toni, T., Welch, D., Strelkowa, N., Ipsen, A. and Stumpf, M.P., 2009. Approximate Bayesian computation scheme for parameter inference and model selection in dynamical systems. Journal of the Royal Society Interface, 6(31), pp.187-202.

    \bibitem{sensitivity} Gutenkunst, R.N., Waterfall, J.J., Casey, F.P., Brown, K.S., Myers, C.R. and Sethna, J.P., 2007. Universally sloppy parameter sensitivities in systems biology models. PLoS Comput Biol, 3(10), p.e189.

\end{thebibliography}

\end{document}