### 2020-08-09 MSc-dissertation.tex
Bib restructure

\documentclass[12pt,a4paper]{report}

\usepackage{epcc}
\usepackage{graphics}
\usepackage{hyperref}
% \usepackage{mathtools}
\usepackage{amsmath}
\usepackage{float}

\begin{document}

\pagenumbering{roman}

\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}

\author{Chaolin Han}

\date{\today}

\makeEPCCtitle

\thispagestyle{empty}

\vspace{11cm}

\begin{center}

    \large{MSc in High Performance Computing}

    \large{The University of Edinburgh}

    \large{Year of Presentation: 2020}

\end{center}

\newpage

\begin{abstract}
    This is the bit where you summarise what is in your thesis.
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listoftables
\listoffigures

% \begin{titlepage}
%     \vspace*{2in}
%     % an acknowledgements section is completely optional but if you decide
%     % not to include it you should still include an empty {titlepage}
%     % environment as this initialises things like section and page numbering.
%     \section*{Acknowledgements}

%     This template is a slightly modified version of the one developed by
%     Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
%     acknowledgement follows:

%     {\em This template has been produced with help from many former students who
%     have shown different ways of doing things. Please make suggestions for
%     further improvements.}

% \end{titlepage}

\pagenumbering{arabic}

\input{introduction/introduction.tex}

\input{background/background.tex}

\input{modelling/modelling.tex}

\input{implementation/implementation.tex}

\input{performance/performance.tex}

\input{conclusion/future_works.tex}

\input{conclusion/conclusion.tex}

% \chapter{Results and Discussions}

% [WHAT results is outputted and WHAT topics are discussed]

% \section{ABC SMC results}

% [inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]

% [model selections result; bayes factor]

% \section{Discussions}

% [goodness of fit (evaluation); preferred models and effects of modification to basic model]

% [efficiency and trade-offs in scaling-up]

% [POSSIBLE: compare to exact inference; compared to other implementations]

% [limitations]

\input{appendix/appendix.tex}

\bibliographystyle{ieeetr}

\bibliography{bibtex/ref}

\end{document}