### 2020-08-17 MSc-dissertation.tex
Fig fix

\documentclass[12pt,a4paper]{report}

\usepackage{epcc}
\usepackage{graphics}
\usepackage{hyperref}
% \usepackage{mathtools}
\usepackage{amsmath}
\usepackage{float}

\begin{document}

\pagenumbering{roman}

\title{Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models}

\author{Chaolin Han}

\date{\today}

\makeEPCCtitle

\thispagestyle{empty}

\vspace{11cm}

\begin{center}

    \large{MSc in High Performance Computing}

    \large{The University of Edinburgh}

    \large{Year of Presentation: 2020}

\end{center}

\newpage

\begin{abstract}
    Zebrafish can regenerate the spinal cord after injury, and possibly recover swimming function. The regeneration is a complex biological process with many interacting parts. We modelled the dynamics of two cell and two molecules that participate in the regeneration, according to the existing experimental findings\cite{ref:Tsarouchas}, and wished to explore uncertainties in their parameter estimates.

    As an Approximate Bayesian Computation (ABC) method, ABC SMC (Sequential Monte-Carlo) method was implemented on the proposed models to approximate the posteriors the parameters. We conducted some preliminary experiments to study the ability of inference and robustness of the algorithm before we performed the inference of parameters and model comparison using the same inference framework. The result shows well-inferred parameters and a preferred model with high model probability.

    The algorithm implementation shows a decent scaling-up performance; however, problems with local optimum and some other uncertainties were also found. Suggestions on the implementation options and ABC SMC settings were made to avoid these problems.
\end{abstract}

\pagenumbering{roman}

\tableofcontents
\listoftables
\listoffigures

% \begin{titlepage}
%     \vspace*{2in}
%     % an acknowledgements section is completely optional but if you decide
%     % not to include it you should still include an empty {titlepage}
%     % environment as this initialises things like section and page numbering.
%     \section*{Acknowledgements}

%     This template is a slightly modified version of the one developed by
%     Prof. Charles Duncan for MSc students in the Dept. of Meteorology. His
%     acknowledgement follows:

%     {\em This template has been produced with help from many former students who
%     have shown different ways of doing things. Please make suggestions for
%     further improvements.}

% \end{titlepage}

\newpage

\pagenumbering{arabic}

\input{introduction/introduction.tex}

\input{background/background.tex}

\input{modelling/modelling.tex}

\input{implementation/implementation.tex}

\input{performance/performance.tex}

\input{conclusion/conclusion.tex}

% \chapter{Results and Discussions}

% [WHAT results is outputted and WHAT topics are discussed]

% \section{ABC SMC results}

% [inferred parameters: joint distribution, estimated values and simulated trajectory for each models; sensitivity; features of the inferred model]

% [model selections result; bayes factor]

% \section{Discussions}

% [goodness of fit (evaluation); preferred models and effects of modification to basic model]

% [efficiency and trade-offs in scaling-up]

% [POSSIBLE: compare to exact inference; compared to other implementations]

% [limitations]

\input{appendix/appendix.tex}

\bibliographystyle{ieeetr}

\bibliography{bibtex/ref}

\end{document}