### 2020-08-18 appendix/appendix.tex
Big fix

\appendix
% the appendix command just changes heading styles for appendices.

\chapter{System and software environment}

\section{ABC SMC implementation}

\subsection{Local machine}

Local development machine is a Mac laptop, running on macOS 10.15.6. The environment of the development is listed in Table \ref{table:local_macine}.

\begin{table}[H]
    \centering
    \begin{tabular}{|c c|}
        \hline
        Environment        & Version                \\ [0.5ex]
        \hline\hline
        Operating system   & macOS Catalina 10.15.6 \\
        PyCharm (IDE)      & Professional 2020.2    \\
        Python interpretor & 3.7.0                  \\
        IPython            & 7.12.0                 \\
        Clang              & 6.0 (clang-600.0.57)   \\
        \hline
    \end{tabular}
    \caption{Environment on local machine}
    \label{table:local_macine}
\end{table}

Version requirements for some critical site packages (\verb|Python|) used in the code are listed in Table \ref{table:local_package}. When replicating the experiments, higher version of these packages are acceptable, except \verb|bokeh|.

\begin{table}
    \centering
    \begin{tabular}{|c c|}
        \hline
        Environment & Version \\ [0.5ex]
        \hline\hline
        pyABC       & 0.10.3  \\
        NumPy       & 1.18.4  \\
        SciPy       & 1.4.1   \\
        pandas      & 1.1.0   \\
        matplotlib  & 3.0.1   \\
        bokeh       & 1.4.0   \\
        \hline
    \end{tabular}
    \caption{Site packages on local machine}
    \label{table:local_package}
\end{table}

Local development and some runs of small size were done under the hardware listed in Table \ref{table:local_hardware}. When running ABC SMC, the program can make used of 8 python process (Hyper-Threading enabled) and run under 3.8GHz (maximal) for a long time under proper cooling, proved that the program could efficiently use the computation resources for a local personal computer. The execution time and other performance data presented in Chapter 5 were obtained using remote machines but not local machine.

\begin{table}
    \centering
    \begin{tabular}{|c c|}
        \hline
        Hardware        & Detail                                \\ [0.5ex]
        \hline\hline
        CPU             & Quad-Core Intel Core i5 8259U 2.3 GHz \\
        Architecture    & 64 Bit                                \\
        Hyper-Threading & Supported                             \\
        Turbo Boost     & Max 3.8 GHz                           \\
        Memory          & 16 GB                                 \\
        Graphics card   & Iris Plus Graphics 655                \\
        \hline
    \end{tabular}
    \caption{Hardwares on local machine}
    \label{table:local_hardware}
\end{table}

\subsection{Remote machine}

Most experiments were performaed on Cirrus, its hardware can be found in Table \ref{table:remote_hardware}. The program can run in sigle compute-node which contains two processors\footnote{\url{https://www.cirrus.ac.uk/about/hardware.html}}. The software environment used in the development and experiments are listed in Table \ref{table:cirrus_soft}. Requirement of site packages is the same as that on local macine (Table \ref{table:local_package})

\begin{table}
    \centering
    \begin{tabular}{|c c|}
        \hline
        Hardware        & Detail                                          \\ [0.5ex]
        \hline\hline
        Processors      & 2.1 GHz, 18-core Intel Xeon E5-2695 (Broadwell) \\
        Architecture    & x86-64                                          \\
        Hyper-Threading & Supported                                       \\
        Memory          & 256 GB NUMA, shared between two processors         \\
        \hline
    \end{tabular}
    \caption{Cirrus hardware (compute node)}
    \label{table:remote_hardware}
\end{table}

\begin{table}[H]
    \centering
    \begin{tabular}{|c c|}
        \hline
        Environment           & Version                      \\ [0.5ex]
        \hline\hline
        Operating system      & Red Hat Enterprise Linux 8.1 \\
        miniconda environment & 4.8.3                        \\
        Python interpreter    & 3.7.7                        \\
        gcc                   & 6.3.0                        \\
        \hline
    \end{tabular}
    \caption{Environment on local machine}
    \label{table:cirrus_soft}
\end{table}

% \section{Data analysis}

\chapter{Supplementary figures and data}

% \section{Infer-back experiments}

% \subsection{Parameter values used to generate synthetic data}

\begin{table}[H]
    \centering
    \begin{tabular}{|c c|}
        \hline
        Parameter            & Value \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 2.20  \\
        $\kappa_{N\beta}$    & 3.96  \\
        $\mu_N$              & 1.72  \\
        $\nu_{N\Phi}$        & 0.219 \\
        \hline
        $\lambda_\Phi$       & 1.31  \\
        $\kappa_{\Phi\beta}$ & 0.124 \\
        $\mu_\Phi$           & 0.145 \\
        \hline
        $s_{\beta N}$        & 6.55  \\
        $i_{\beta\Phi}$      & 1.71  \\
        $\mu_\beta$          & 0.521 \\
        \hline
        $s_{\alpha\Phi}$     & 10.2  \\
        $\mu_\alpha$         & 19.7  \\
        \hline
    \end{tabular}
    \caption[Parameter values used to generate synthetic data for model 1]
    {Parameter values used to generate synthetic data for model 1. They were obtained from a preliminarily least-square fitting}
    \label{table:known_values}
\end{table}

% \subsection{Kernel experiment: median epsilon schedule}

\begin{figure}[H]

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel2.png}}
    \end{center}

    \caption[Total sampling size of different kernels, using median epsilon strategy]
    {Total sampling size of different kernels, using median epsilon strategy. Different color represents different generations (bottom to top: population 1 to population 20)}
    \label{fig:kernel2}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance2.png}}
    \end{center}

    \caption[Acceptance rates of different kernels, using median epsilon strategy]
    {Acceptance rates of different kernels, using median epsilon strategy. Each population has 2000 particles. The total length of populations is different because some runs reached the final $\epsilon_T$ earlier than others}
    \label{fig:acceptance2}

\end{figure}

% \section{Parameter inference and model comparison results}

\begin{figure}[H]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/credible.png}}
    \end{center}
    \caption[Credible intervals of parameters in model 5]%
    {Credible intervals of parameters in model 5 across all generations}
    \label{fig:credible}
\end{figure}