### 2020-08-20 implementation/implementation.tex

\chapter{Parameter estimation and model comparison}

For the proposed models, exploration of the uncertainties in the parameter estimations is our main task in this chapter, which involved the application of ABC SMC and several experiments. After that, the parameter inference framework was also used for model selection.

The build and test of the code were under \verb|Python| environment with \verb|pyABC|\cite{ref:pyabc}. Some other code, e.g. shell scripts and \verb|R| code were also used to perform the experiments and result analysis. Software and system environment used in our implementation can be found in Appendix A. The build and test were performed on local computers and HPC facilities available within EPCC.

\section{Code and implementations}

%  [ABC implementations details, e.g. distance, population, ODE related functions]

%  [how the code is developed and built]

%  which is of our interests and suitable for the followed performance experiments.

According to our preliminary studies, an ABC SMC framework implemented using \verb|pyABC| in \verb|Python| was adopted in this project. \verb|pyABC| is a popular SMC software packages\cite{ref:pyabc} which has been used form many ABC SMC inference applications\cite{population}. \verb|pyABC| provides an open-source framework for likelihood-free inference, which is a \verb|NumPy| implementation of ABC SMC algorithm and a useful tool-box for our inference and analysis tasks. Besides, it supports multi-core execution implemented using \verb|multiprocessing| built-in package, and it made our further studies in the scaling-up performance possible.

The inference framework code for the project was developed and tested in a local environment (macOS 10.15.6, see Table \ref{table:local_macine}) at first using a small input, then the functional version was deployed on compute node of ARCHER and Cirrus for parameter inference experiments and model selection tasks. The scaling-up experiment was performed on Cirrus. 
%The profiling and its analysis were performed on a local machine. 
Some necessary development and debug were performed on the remote machines to ensure that experiments run correctly on the compute node.

The ODE model-related code and utilities were first developed to enable reusable functions, variables and data calls. It included the code for the ODE equations, ODE solver, data structures and data format transform functions, etc. By doing this modularisation in separate code files, we could change implementation options and activate ABC SMC easily in another `trigger' file, without creating definition and duplicated code for the models, observed data and parameter priors.

Some missing features were found when using \verb|pyABC|; therefore some modifications were made to the source code of \verb|pyABC|.For example, the laboratory-measured data were not considered `complete' or `tidy' for modelling and analysis purpose: the available time points for cells and cytokines were different, e.g. cell number was not measured at 0.25 hpl and cytokines' expression was not measured at time point 120 hpl. These missing values were treated as `ignored' when calculating the distance between observed data and simulated data, and represented by \verb|NaN|s. However, in the built-in distance calculations, \verb|NaN| was not acceptable. To cope with this, the built-in distance function was modified. Other modified code included some visualisation and result representation enhancements.

Analysis of the program output was also wrapped up into functions which can be easily called after an ABC SMC run is finished, which include multiple visualisations and out of summary statistics of the results. For different tasks, there are also separated analysis code file. Other analysis tools used in the project include built-in database visualisation tool, R code and Microsoft Excel.

The project code was managed via \verb|git| version-control and available as a GitHub repository\footnote{\url{https://github.com/chaolinhan/MSc_Project}}; it also made the cross-hardware synchronisation of code versions convenient.

After several development iterations, code files for ABC SMC implementation now mainly had two types: code files for infer-back experiments (using synthetic data), and code files for parameter inference (using experimental data). Model selection can be easily performed by adding more model to the implementation code file.

% \section{Parameter estimation}

% [implementation details and options/ settings that are affecting the ABC]

\section{Hyperparameters and implementation options experiments}

As the key focus of this project was on the parameter estimations of a dynamic system, the ability of inference on the proposed models was firstly examined before actual inference using the experimental data.

In the first part, synthetic data (Table \ref{table:known_values}) was used with known true parameter values. The algorithm implementation was evaluated by the efficiency and goodness of the resultant model under different implementation options and ABC SMC hyperparameters.

Then according to findings in the first part, several options with considerable performance were tried in the actual inference with observed data (Figure \ref{fig:obs_data}). To obtain more accurate and general results, these experiments were repeated for three or more times.

    % [including the options that are tested/ tried/ not adopted]

    % [including experiments results and discussions]

    % [how do these options correspond to biological model and how to set them accordingly for proposed model]

    % [why use synthetic data]

    The ABC SMC inference can have widely different performance when using different hyperparameters and implementation options. To firstly test the ability of inference and observe the results of different options, synthetic trajectories with known true parameter values were used as target data. By doing this, we could directly compare the inferred parameters to the true values and compare the simulated trajectories to the true trajectories, then observe the result to see whether the inference is successful and how well the model fits the data. Also, by setting the same final threshold value, the efficiency under different options can be compared and help us to choose proper options when applying ABC SMC.

    The true values (Table \ref{table:known_values}) were obtained by fitting model 1 onto the experimental data (Figure \ref{fig:infer_back_data}) using least-square optimisation method. Available least-square fit tools in \verb|SciPy| were used. The fitting algorithm was Trust Region Reflective\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.optimize.least_squares.html}}. Target data was then generated using these parameter values via ODE integrate solver\footnote{\url{https://docs.scipy.org/doc/scipy/reference/generated/scipy.integrate.odeint.html}}.

\begin{figure}[ht]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/LS.png}}
    \end{center}

    \caption[Synthetic data generated with known parameter values]%
    {Synthetic data generated with known parameter values (line plot), compared to experimental data (scatter plot). Known parameter values are obtained from a least square fitting of the observed data. Error bar indicates SEM}
    \label{fig:infer_back_data}

\end{figure}

The following topics were studied using the synthetic data and the developed ABC SMC inference framework.

\subsection{Perturbation kernels}

Perturbation kernels work in the sampling process. In each generation $t$, samples are firstly taken from the previous population $\{\theta_{t-1}\}$ with weights $\{w_{t-1}\}$, then perturbed using the perturbation kernel $K(\theta|\theta^*)$. We keep sampling until $N$ particles are accepted, where $N$ is the pre-set population size. After that, the new weights are calculated and normalised. Figure \ref{fig:smc} illustrates the process.

The perturbation kernel is called in the sampling of every particle and determines how the new perturbed particle is chosen; thus it is influential to both computation complexity and the resulting posterior distribution. Generally, a local perturbation kernel may face the risk of being stuck in local modes (e,g., local optima), but it may need less computational operations or could generate a population with a higher acceptance rate. A kernel with a wide variance, or spreading out in a large space could help in resolving the local optima problem by a more thorough exploration of the parameter space. However, it can be more computation-intensive and result in lower acceptance rates. A desired kernel should balance the trade-offs; their properties and criteria were discussed in \cite{ref:kernel}.

There are several common choices of perturbation kernels. Among them, the multivariate normal kernel and local M-nearest neighbour model is preferred to be applied to our models. A covariance matrix $\Sigma_t$ of accepted particles is calculated from the previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$. It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in reflecting the true posterior structure \cite{ref:kernel}. It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks. The $scaling\in(0,1]$ option in \verb|pyABC| will be multiplied to the covariance to produce a `thinner-distributed' perturbation result.

Local M-NN kernel provided by \verb|pyABC| was also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.

The kernel experiment was designed to explore the efficiency of SMC on our dynamical systems. Given the same fixed threshold schedule, kernels that need less number of samples and have higher acceptance rates would be our preference. The experiment compared multivariate normal kernel with local MNN kernels with different options, using synthetic data to infer back the parameters. The acceptance rate among each generation and total required number of sample were compared to give suggestions on the kernels. As the threshold schedule was fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit (i.e. recover of target data trends/features and errors of the inferred parameters compared to true values) were not discussed.

\subsubsection{Kernel experiment results}

To compare kernels, a fixed schedule of thresholds was used with the final epsilon $\epsilon_t=10.0$. From the total required number of samples plot (Figure \ref{fig:kernel1}) the efficiency can be compared. Among the tested kernels, the local M-NN with M=50 has the best performance: it requires the least number of samples and
has higher acceptance rates among almost all generations. Local M-NN with M=750 is the slowest kernel.

For local M-NN kernels, generally, a higher M leads to lower acceptance rates and more required particles in each generation. As shown in Figure \ref{fig:acceptance1}, local M-NN with M=750 is the lowest curve. Consequently, if a greater M was tested, then it would have even lower acceptance rates; the maximum M=2000 (the whole population is considered) would have the lowest acceptance rates.

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/kernel1.png}}
    \end{center}

    \caption[Total required number of samples for different kernels]%
    {Total required number of samples of different kernels after 20 populations (2000 particles in each population). Different color represents different generations (bottom to top: population 1 to population 20)}
    \label{fig:kernel1}

    \vspace*{\floatsep}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
    \end{center}

    \caption[Acceptance rates of different kernels]%
    {Acceptance rates of different kernels in among populations. Each population has 2000 particles}
    \label{fig:acceptance1}

\end{figure}

Multivariate normal kernel has a performance between local M-NN M=250 and M=100. It proves that rather than a trivial normal kernel, multivariate normal kernels are more efficient facing the concentrations of joint distributions among multiple parameters \cite{ref:kernel}. `Scaling' option can narrow the distribution calculated from the last population, thus making the kernel more `local' by sampling under a smaller variance. Similar to small M in local M-NN, small `scaling' is also more efficient in our experiment of model 1. Besides the fixed threshold schedule, a median threshold schedule with the same final threshold $\epsilon_{20}=10.0$ was also tested and similar results are observed (Figure \ref{fig:kernel2} and \ref{fig:acceptance2}).

A more `local' kernel is more likely to give better performance by efficiently sampling around concentrations in parameters' distribution and approximate the posterior distribution more quickly. This holds in most cases under a given target final threshold value $\epsilon_t$. However, it may not produce the proper result as  we want: a more local kernel, e.g. local M-NN with small M or multivariate normal with `scaling' is more likely to be stuck in local optima, as the kernel is less likely to sample particles that are far from local concentrations, although these local optima are still acceptable before even smaller threshold values is demanded (see Section 5.2 for more discussions). One obvious example is presented in Figure \ref{fig:kernel1}, where local M-NN with M=50 suffered from a local optimum: the last generation takes much more samples to meet the required threshold.

% In other words, by using local kernels the epsilon may quickly converge to a local optimum; if an even smaller epsilon is desired, the local kernel can hardly generate enough particles to find another matched local modes.

% Also, the shapes of posterior distribution affect the performance of local kernels \cite{ref:kernel}.

To conclude, Local kernels can be fast but unstable in some case. For a general parameter inference for our models, multivariate kernels are preferred, as a good fit is our prior target. Some local kernels are also worth trying after the multivariate normal kernel, e.g. local M-NN with M$\leq 250$ (for a population size of 2000, i.e. 12.5\% of the population) and multivariate normal with scaling, but extra cares must be taken: although they are fast, moving in and moving out of local optimum make them unstable in efficiency.

% Also, multivariate M-NN is worth trying, although it was not built in \verb|pyABC| and requires additional implementations.

% \begin{figure}
%     \begin{center}
%     \resizebox{1.0\hsize}{!}{\includegraphics{fig/acceptance1.png}}
%     \end{center}

%     \caption{Acceptance rates of different kernels}
%     \label{fig:acceptance1}

% \end{figure}

\subsection{Adaptive functions and factors}

% [how adaptive distance work]

SMC can be more adaptive by introducing adaptive distance function and adaptive population; besides, factors can be applied to manually `normalise' the data.

The trivial distance function is set to Euclidean distance (2-norm distance)

\begin{align}
    \label{eq:dis}
    D=\sqrt{\sum_i \Delta x_i^2}
\end{align}

where $i$ is the index of data points and $\Delta x_i$ is the discrepancy between  observed data and simulated data at data point $i$, i.e. $\Delta x_i = x_{i, simulated}-x_{i, observed}$. In this case, data points are 12 time points for four variables, i.e. 48 data points in total.

A weighted 2-norm distance can be written as

\begin{align}
    \label{dis_w}
    D=\sqrt{\sum_i w_i \Delta x_i^2}
\end{align}

where a weight $w_i$ is assigned to every data point. If data point $i$ has a higher weight, then the data value is regeared to be more informative, i.e. giving more help in inferring the true posteriors. Weights can be either pre-set according to prior knowledge of the problem or updated using an adaptive method. The later method is known as adaptive distance.

Adaptive distance is changing the weights of all data points after each generation iteration, trying to assign informative data points higher weights. Additionally, implementation of adaptive in \verb|pyABC| also introduces additional factors $f_i$ multiplied to weights \cite{ref:adpt_dis}.

\begin{align}
    \label{dis_f}
    D=\sqrt{\sum_i f_iw_i \Delta x_i^2}
\end{align}

Factor is helpful as an option for efficiency. When some data points are equally informative, the manually pre-defined factor can make the distance more focuse on certain data points, or behave like a normalisation that can balance the scales of different summarised statistics. There are two appliance case studied: (1) factors used as a normalisation option and (2) factors used to give more focus on some features of the observed data. For (2), as we noticed that the main features (e.g. rapid increase, peaks, fluctuations) are mostly observed in the first half of the time points, i.e. 0 - 30 hpl. Such factor scheme assigns data in the first half of the time point with higher factors (75) and data in the second half with lower factors (25).

Adaptive population \cite{population} can adapt the population size of each generation according to the mean correlation of variation error of the previous population. However, in our early tests,  the maximal allowed population size is set to 10,000, and the adaptive strategy always selected the upper bound (10,000) for new populations, as a result of wide parameter space or the wide posterior approximation shape. Additionally, some internal bugs were observed when applying it. Hence, the adaptive population was not tried in this project; however, it is worth trying in future works as an efficiency improvement.

% [what is factor]

% [what is adaptive distance]

\subsubsection{Experiment results}

% [FIGURE here]

Two types of factor and adaptive distance function were tested by running the same number of generations (20 generations, 2000 particles per generation). The first type of factor is 75:25 factor, where the first half of the trajectory data were assigned higher factor (75), and the second half of the data was assigned lower factor (25); this could make the first half of the data more `important' in the inference. The second factor was `range' factor, which tried to normalise the trajectory of each dependent variable according to its data range.

From Figure \ref{fig:factor}, different types of factor gave different performance. 75:25 factor required much more samples in the first generation (orange bar in the bottom of 25:75 factor plot, left of Figure \ref{fig:factor}) and also did not improve much in later generations; thus, it is the slowest one. The first half of observed data contained more features and information, assigning it with higher factor would make the algorithm think that it is more important to fit the first half and thus focus on the hard part, and consequently more particles were needed in the inference process. The acceptance rates of 75:25 factor among different generations were also generally slightly lower than the standard. By contrast, range factor required less total samples and gave generally higher acceptance rates than standard, which made it considerably faster. The adaptive distance gave the best efficiency, as it significantly improved the acceptance rates and resulted in even less required samples (right, Figure \ref{fig:factor}).

\begin{figure}[t!]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/ib.png}}
    \end{center}

    \caption[Factors and adaptive distance experiments]%
    {Factors and adaptive distance experiments. Left: total required number of samples. Different color represents different generations (bottom to top: population 1 to population 20). Right: acceptance rates in each generation. }
    \label{fig:factor}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/ib_factor.png}}
    \end{center}

    \caption[Simulated trajectories ]%
    {Simulated trajectories from the last population in the factor and adaptive distance experiment. 500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from. 75:25 factor showed very similar curve as the range factor.  
    }
    \label{fig:factor_sim}

\end{figure}

However, the result of factors adaptive distance could not be directly compared to standard run, because it changed the distance measurement by adapting weights or manually setting factors. Factors were normalised to make it closer to the standard distance measurement. As a result, the same final threshold value did not guarantee that adaptive distance can reach the same level of fit as others. After 20 populations, adaptive distance had a poor fit (Figure \ref{fig:factor_sim}), while the standard run and two types of factors had similar tight fits of the data, which all converge to the true posterior. Additional in our implementations, variance factor (which tries to normalise the data according to the dependent variables' variance) was also tried and gave a very similar performance as the range factor.

To conclude, because the factors and weights are directly multiplied to $\Delta x$, and thus the metrics for distance are changed, the direct comparison of the efficiency under the same threshold schedule was unfeasible; we could only compare the results after the same number of generations. Factors gave nearly the same inferred results but required different numbers of samples, hence cares must be taken when choosing factor schemes. Adaptive distance did not provide an accurate result as others. As a result, our further implementations used standard distance function, and factors were also tried for possible improvements.

% by applying factors the total required samples are less, and the adaptive distance requires much less samples; their acceptance rates are also higher than the standard one. It seems that adaptive function and factors could help to converge more quickly.

% However considering the result after 20 generation (Figure \ref{fig:factor_sim}), the resultant model gives the contrary preference. Applying factors and using adaptive distance can lead the resultant model less accurate (after the same generation iterations): they have wider inter-quantile range and may requires more generations to converge. TODO

% \subsubsection{Conclusion} We noticed the significant efficiency improvement of these adaptive options; they can largely improve the acceptance rate in some cases. However, as the factors and weights are directly multiplied to $\Delta x$ and thus the metrics for distance are changed, a direct compare of the efficiency under the same threshold schedule is unfeasible; after the same number of generations they can reach a small $\epsilon_t$ but still the approximated posterior is not as accurate as the standard implementation. AS a result, our implementation of the parameter inference will only consider standard distance function and factors are only tested for improvements (FIGURE).

\subsection{Data size, prior distribution and population}

% [experiments plan]

Some other hyperparameters, e.g. feed-in data size, prior distribution, number of populations and number of particles in each population were also of our interest. Experiments were designed to explore how these options affect the goodness of fit and efficiency of the implementation.

Regarding the existing dynamic systems model, SMC was applied with two data size options and two ranges of the prior interval. The population options, i.e. population size and the number of populations were also tried with different values. These experiments intended to give suggestions on the later implementation on experimental data.

\subsubsection{Experiments results}

\begin{figure}[H]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/size1.png}}
    \end{center}

    \caption[Data size and prior distribution range experiment]%
    {Data size and prior distribution range experiment. Left: total required number of samples. Right: epsilon values under median epsilon schedule}
    \label{fig:size}

\end{figure}

The results from data size and prior distribution range are shown in Figure \ref{fig:size}. The standard implementation is fed with 120 data points (30 data points for each of the four variables), the `less data' is fed with 48 data points, which is the case of the real experimental measurement. Although much less (60\%) data was used, the total required sampling numbers was only 4.47\% less. This indicated that the data size does not affect the sampling process much and the inferred model are all considered acceptable when compared to the synthetic data, as they finally reached a similar epsilon threshold (Figure \ref{fig:size}), but cares should still be taken when using less data points, or using other more summative statistics (e.g. mean, standard deviation, peak values, starting and ending values etc.): as shown in \cite{ref:disease}, fewer data can lead to a model with more variance, under-fitting or missing some local features such as local peaks.

The prior range and distribution can largely affect the inference progress (Figure \ref{fig:size}). Samples were drawn from a high dimensional parameter space in each sampling process. A wider parameter prior interval in only one dimension can cause the parameter space to be much bigger. Consequently, the exploration and sampling of parameter vectors would take much more efforts. Right plot in Figure \ref{fig:size} also indicates that the convergency is slower for wider prior ranges. As a result, a proper prior interval based on our experience and belief without seeing the data should be set as narrow as possible, as the inference is largely affected by the prior interval and its length. Improper prior could even make the algorithm miss the true posterior \cite{ref:abcsysbio}.

How to specify prior can be tricky from problem to problem. For our task in the parameter inference of regeneration model, biologically motivated priors were used, which were a general reflection of biological constraints. Also, the units of the parameters helped us in identifying the possible scales of the parameters.

% taken from a high dimensional parameter space in each sampling process, wider prior distribution ranges will making each sampling explore more in the parameter vector space. It suggests that a narrower prior is preferred and we should make a more accurate and confident prior belief as possible for the efficiency concerns.

%     [discuss uniform and log uniform when result ready]

% [FIGURE HERE]

\subsubsection{Choise of options}

This section aims to explore options in ABC SMC implementations preliminarily. From the above results, we adopted multivariate normal kernel as the kernel for our real inference on observed data; Local M-NN kernels were also planned to be tested. In the real inference, 75:25 factor were also tried to see if it could help in recovering more features and resulting in a better fit. Prior intervals were explored explicitly and gradually shrunk for our models.

% Because the improvement by using less data was inapparent, we did not try to change the target summary data size.

Some other hyperparameters, e.g. the number of generations, population size (i.e. the number of particles per generation) were not addressed in this subsection. In the real inference, population size was set to be greater than 2000 to avoid insufficient exploration of the previous population; in some case than high variance or uncertainties in the results were observed, as large as 20,000 were used as population size. If the epsilon did not show a convergency trend, the run could be extended to more generations until a stable result was observed.

% using synthetic data with known true parameter values, thus the efficiency and the goodness of the resultant model can be compared across these options. Conclusion can be drawn to choose efficient and proper options of ABC SMC to be applied on the target models with experimental data. It also provides reference for a general inference task using SMC with high dimensional parameter space.

\section{Parameter estimation and model comparison}

\subsection{Model 1, 2 and 3}

% [separated ABC run on model 1, 2 and 3]

Using the suggested options in the previous section, the target data (Figure \ref{fig:obs_data}) was prepared as input for the SMC inference framework. Regarding the result, the population size and number of generations were adjusted several times to obtain a general informative posterior which should be stable across repeated runs and avoid local optima as much as possible.

Two prior intervals ([0, 25] and [0, 75]) and two distribution shape (uniform and log-uniform) were tried in the parameter estimation of the first three models. All these ABC SMC runs used 2,000 particles per population and 30 generations. The population size should be set at a high level (e.g. $\geq 1000$ in our case) when having models with many parameters to ensure a statistically adequate exploration. The convergency were usually observed around generation 20 in our previous experiments; the length of generation is set to 30 to ensure that most runs converge in the end.

In practice, we found that log-uniform-shaped prior gave a better fit of the model than uniform-shaped prior, and wider prior range [0, 75] does not give a more accurate result than [0, 25]. Then prior interval [0, 50] was tried, in case some parameters have a true posterior laying between [25, 50].

The results of prior distribution log-uniform [0, 50] is shown in Figure \ref{fig:result123}, which is considered to give a better fit than other prior options. For these models, the acceptance rates and epsilon path are presented in Figure \ref{fig:result123_2}, and the estimated parameters' posterior of model 3 is shown in Figure \ref{fig:para1}. Figure \ref{fig:result123} plots the simulated trajectories of the three models. One thousand particles (each particle is a parameter set) were sampled for the last generation, and the mean (blue) and 25th to 75th percentile interval (grey) were calculated from the data simulated by these 1000 particles.

Before applying model comparison methods, some features of the three model can still be compared. From simulated curves, It can be seen that all the inter-quartile ranges are tight around the mean value curve, which indicates the approximated posterior are concentrated; the bell-shaped posterior distributions with a single peak (Figure \ref{fig:para1}) agrees with this. All epsilon values converge to a steady level, and model 3 has a lower final epsilon value, and consequently, its simulated trajectory is more close to the observed data (model 3, Figure \ref{fig:result123}).

The acceptance rates are fluctuating and have a gradually increasing trend for all the three models, as the acceptance rate becomes higher when the true posteriors are gradually approximated. Compared to model 1, model 2 gives a better fit for the decreasing trend in the second half of the observed data. This proves that using an exponentially decaying self-increase rate $\lambda_N$ instead of a constant term indeed help to improve the model.

The resultant simulated data from model 2 and 3 have similar trends, while the most obvious difference is that for model 3, the simulated data of $\Phi$ and $\alpha$ are more `flat'. We expected model 3 to be the best as it reached a lower threshold value, and intuitively gives a better fit. However, the fit of all these models could not be considered as a generally good representation of the biological process that we want to approximate: for tnf-$\alpha$, the simulated data and the observed data show largely different trends. A sharp peak at 4 hpl was observed in the experimental data, but none of the three models represents similar features (Figure \ref{fig:result123}).

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve123.png}}
    \end{center}

    \caption[Simulated trajectory from the 20th population of model 1, 2 and 3]%
    {Simulated trajectory from the last population of model 1, 2 and 3. Error bar indicates SEM. 500 randomly chosen particles from last population are used to generate a sequence of trajectories where mean and inter-quartile range are calculated from}
    \label{fig:result123}

    \vspace*{\floatsep}

    \begin{center}
        \resizebox{0.6\hsize}{!}{\includegraphics{fig/eps_acc123.png}}
    \end{center}

    \caption{Epsilon trends and acceptance rates of model 1, 2 and 3}
    \label{fig:result123_2}

\end{figure}

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/para1234.png}}
    \end{center}

    \caption[Estimated posterior distribution of parameters in model 3]%
    {Estimated posterior distribution of parameters in model 3. Shaded range indicates inter-quartile interval}
    \label{fig:para1}

    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/cmp1.png}}
    \end{center}

    \caption[ABC SMC based model comparison of model 1, 2 and 3]%
    {ABC SMC based model comparison of model 1, 2 and 3. From population 8 to population 30, model3 always wins}
    \label{fig:cmp1}

\end{figure}

% [model comparison among model 1, 2, 3]

Next, a model selection experiment with different prior was conducted, using the same ABC-SMC framework. As in the parameter inference, two distributions (uniform and log-uniform) are tested with three different interval ([0, 25], [0, 50] and [0, 75]). The results show that model 3 wins with 100\% model probability after 30 generations for most runs, except log-uniform [0, 75]. All the model probability results are of great discrepancy (e.g. 100\% for model 3 and 0\% for model 1 and 2), so Bayesian factor is not calculated.

Figure \ref{fig:cmp1} shows the model selection process, under log-uniform [0, 50] prior. Model 3 gains total advantage over model 1 and 2 after the seventh population. Together with the conclusions above, model 3 is considered to be the best model to represent the target data so far. Considering the motivations in proposing model 2 and 3, the result shows that a decaying $\lambda$ term is reasonable in both theory and practice; simplification of $i_{\beta\Phi}$ term gives an even better model (i.e. model 3), which means that the inhibition effect from macrophage can indeed be ignored.

\begin{table}[t!]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Estimated mean  & Estimated mean      \\
                             & value (uniform) & value (log-uniform) \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 14.1            & 13.3                \\
        $a$                  & 9.67            & 0.0170              \\
        $\kappa_{N\beta}$    & 18.0            & 0.259               \\
        $\mu_N$              & 13.2            & 0.0180              \\
        $\nu_{N\Phi}$        & 0.742           & 0.131               \\
        \hline
        $\kappa_{\Phi\beta}$ & 0.239           & 0.156               \\
        $\mu_\Phi$           & 0.154           & 0.0350              \\
        \hline
        $s_{\beta N}$        & 6.75            & 1.21                \\
        $\mu_\beta$          & 6.38            & 1.32                \\
        \hline
        $s_{\alpha\Phi}$     & 17.0            & 5.83                \\
        $\mu_\alpha$         & 11.9            & 2.43                \\
        \hline
    \end{tabular}
    \caption[Estimated parameter values of model 3]
    {Estimated parameter values of model 3, using uniform prior distribution and log-uniform prior distribution respectively}
    \label{table:estimated1}
\end{table}

Compared to uniform-distributed prior, we found that log-uniform distributed prior is more suitable in our parameter inference. Log-uniform resulted in an obviously better fit and narrower inter-quantile interval (Figure \ref{fig:result123} and \ref{fig:resultCurve_uni}), and the final epsilon value was also smaller. Log-uniform tends to assign values that are close to the left boundary (i.e. zero in our case) with higher density. As a result, the parameter estimates from log-uniform prior are highly likely to have smaller mean values than that from uniform prior, and the inferred parameter values proved this (Table \ref{table:estimated1}). This was also observed in other models and prior interval experiments. It indicated that some of the parameters tend to have small true values that are close to zero, and using log-uniform prior could result in a faster and more satisfying inference. Regarding this, further experiments all used log-uniform prior only.

\subsection{Model 4 and 5}

% [features that derive model 4 and 5]

We observed that the existing models could not well represent the trajectory of relative expression of tnf-$\alpha$. From experimental data, a rapid increase of tnf-$\alpha$ expression was observed in 0--4 hpl, followed by a rapid drop. All the models cannot well recover this feature, and most results saw a less-rapid increase followed by a gradual decrease, while the observed peak value was not reached (Figure \ref{fig:result123}).

Also, in some other experiment results, e.g. uniform runs (Figure \ref{fig:resultCurve_uni}), tnf-$\alpha$ trajectories was well fitted; however, at the cost of under-fitting $\Phi$ curve. Hence efforts had been spent to propose more alternative models that can potentially solve this problem.

\begin{figure}[h]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve_uni.png}}
    \end{center}

    \caption[Simulated data from the last population of model 1, 2 and 3, using uniform prior]{Simulated data from the last population of model 1, 2 and 3, using uniform prior.  Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
    \label{fig:resultCurve_uni}
\end{figure}

The two proposed alternative models are based on model 3 and try to add extra term in the tnf-$\alpha$ equation $\mathrm{d} \alpha/\mathrm{d} t$. As macrophage is the main source of tnf-$\alpha$ production, and a similar sharp increase-and-drop trend is observed in the il-1$\beta$ trajectory, two hypotheses and corresponding terms are propose as follows.

\paragraph{Model 4} It is assumed that the expression of il-1$\beta$ would have a promoting effect, representing by a phenomenological term $d_{\beta\alpha}\beta$ and $d_{\beta\alpha}$ is a new model parameter (positive constant). The promoting effect here is regarded to have the equivalent effect as directly promoting by il-1$\beta$, but the underlying mechanism is unclear so far. In mathematical view, this additional term can accelerate the expression of tnf-$\alpha$ in the first few time points and help to recover the peak-shaped trajectory. The proposed model is written as Equation \ref{eq:model4}.

\paragraph{Model 5} Alternatively, we considered promotions to the tnf-$\alpha$ production, i.e. $s_{\alpha\Phi}\Phi$. It is assumed that il-$\beta$ could accelerate the production process. An additional term $f_{\beta\alpha}\beta$ is introduced to production rate, with $f_{\beta\alpha}$ being a new model parameter (positive constant). This model meets the biological context considering the source of tnf-$\alpha$. The proposed model is written as Equation \ref{eq:model5}.

% [further comparison of model 3, 4 and 5]

Model 4 and 5 are based on model 3, so in this phase experiments of these three models were conducted for separately parameter inferences and overall model comparison. Based on our experience in conducted experiments, we used the following setting for parameter inference:

\begin{itemize}
    \item population size: 2000 and 5000
    \item number of populations: 30 generations
    \item prior: [0, 50] log-uniform
    \item perturbation kernel: multivariate normal kernel
\end{itemize}

Factors (Equation \ref{dis_f}) were also tried to find if we can enforce the inference framework to give more importance to the first few data points. An optional factor scheme was proposed: for each trajectory, the first 8 data points were assigned higher factors (75), and the rest 4 data points are assigned with lower factors (25).

All these experiments were conducted on remote machines, and the output database files were retrieved for analysis.

Figure \ref{fig:resultCurve345} shows the simulated data from the inferred models, with population size 2000. It can be seen that the addressed problem in the trajectory of tnf-$\alpha$ is partially relieved in model 4 and 5: compared to model 3, a peak appears around 4 hpl; it can represent more features in the observed data and consequently the final reached epsilon (after 30 populations) is smaller than that of model 3, which proves that our proposed modifications in model 4 and 5 are helpful in recovering more features in observed data.

\begin{figure}
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/resultCurve345.png}}
    \end{center}

    \caption[Simulated data from the last population of model 3, 4 and 5]{Simulated data from the last population of model 3, 4 and 5. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small)}
    \label{fig:resultCurve345}

\end{figure}

By intuition model 4 and 5 would be the best models so far that can recover most of the observed data features. Model 1, 2 and 3 could also converge to a low final epsilon value (~20), but a key feature observed in target data -- a peak at around 4 hpl -- was not fitted. Some features, e.g. fluctuations in il-1$\beta$ and tnf-$alpha$ trajectories at around 20 hpl were still not ideally fitted. All existing models gave steady or gradually decreasing trends for all the four variables after 20 hpl, where fluctuations in the observed data are fitted by smooth and steady curves.

After applying factors, the results were not largely different. As expected, applying factors made the data points at first half (where exists most of the fluctuations) to be more `focused'. However, the inferred models were not largely different from ones without factors, which indicated that for model 4 and 5, applying factors will not largely affect the final convergency and the resultant posterior.

After we tried more generations with factors, an `over-fitting' behaviour was observed at the trajectory of macrophage (Figure \ref{fig:overfit}). In this case, the discrepancy of simulated data and observed data mainly comes from the last 2 data point (72 hpl and 120 hpl) of macrophage. As additional importance was assigned to the first few data points, the inference algorithm tended to fit first data points preferentially, and thus relatively large discrepancy at the tails were acceptable for it.

\begin{figure}[ht!]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/overfit.png}}
    \end{center}

    \caption[Simulated trajectory from the 30th population of model 5, with 75:25 factor applied]{Simulated trajectory from the 30th population of model 5, with 75:25 factor applied. Error bar indicates SEM (for $N$ and $\Phi$ the SEM is relatively small).}
    \label{fig:overfit}
\end{figure}

Next, we performed a model comparison on model 3, 4 and 5, using the same ABC SMC inference framework. Figure \ref{fig:model345cmp} shows the model probabilities at different generations.

\begin{figure}[ht!]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model345cmp.png}}
    \end{center}

    \caption[Model probabilities of model 3, 4 and 5]{Model probabilities of model 3, 4 and 5 in ABC SMC based model comparison. Model 5 also wins in further populations (population 25 to 30)}
    \label{fig:model345cmp}
\end{figure}

From the model probability plot (Figure \ref{fig:model345cmp}), we found that model 4 did not overweight other models across all the generations. Model 3 was comitative and gained advantages at most early generation. From generation 14 and onwards, model 5 saw a gradually increasing probability and gained 100\% model probability after generation 18 ($\epsilon_{18} = 16.98$) and won in further generations. Also, repeated runs showed similar trends, although uncertainties (e.g. drastically model probability changes) were present before generation 20, model 5 won with nearly 100\% probability in later generations where epsilon values were considerably low, i.e. $\epsilon_t<16$.

Based on the model comparison results, we could conclude that model 5 was the best model among the tested models. 
It proves that the extra term $f_{\beta\alpha}$ not only makes sense regarding the source of tnf-$\alpha$, but also gives a better model than phenomenological term $d_{\beta\alpha}\beta$.
Uncertainties were observed in early generations, where similar patterns were also observed in the model comparison of model 1, 2 and 3. We considered the drastically changing model probability in early generations is related to the inference ability at high epsilon threshold values (epsilon) and possibly the insufficient exploration of the parameter space. However as we only focused on the best model with small final threshold $\epsilon_T$, the easiest way is to extend the number of generation until epsilon converges, or increase the population size for a more statistically confident probability.

% From population 10 to population 20 model 4 gains advantage but after population 20 model 5 is preferred. Again, as DISCUSSED in the model comparison of model 1, 2 and 3, Given the present result, model 5 is regarded as the best model.

% Some further experiments tried larger population size (10,000 particles per population) and more populations (36 populations), but the resultant model did not have a significantly better fit of the observed data, although much more computation resources are used.

For the best model (model 5), the inferred parameter posterior distribution of model 5 is shown in Figure \ref{fig:model5_para}, the estimated mean values is shown in Table \ref{table:estimated5}. The pair-wise joint density distribution is shown in Figure \ref{fig:model5_para_joint}. Most of the inferred posteriors show a bell-shaped density distribution with a single peak.. A thin and concentrated shape indicates a well-inferred parameter, e.g. $\mu_{\Phi}$ and $\kappa_{\Phi\beta}$. Some distributions, however, are skewed to zero, e.g. $\mu_N$ and $s_{\alpha\Phi}$. Concentrations in all the posteriors were presented, and using inter-quartiles (shaded interval in Figure \ref{fig:model5_para}) or mean values to summarise the posteriors is feasible in practice uses, e.g. simulating data (Figure \ref{fig:overfit}).

By looking at the joint density plot (Figure \ref{fig:model5_para_joint}), some correlated patterns are observed. For example, an obvious linear relationship is presented in the joint distribution of $\mu_\beta$ and $s_{\beta\Phi}$. This leads us to consider that we could possibly reduce the number of parameters, e.g. replace $s_{\beta\Phi}$ by $A\mu_\beta+B$ where $A$ and $B$ can be calculated from the joint distribution by fitting a linear regression. The reason for the highly correlated pattern, or `fat' distribution shapes could be that the provided data are lees informative about the parameters.

\begin{figure}[ht]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/model5_para.png}}
    \end{center}

    \caption[Inferred posterior distribution of parameters in model 5]%
    {Inferred posterior distribution of parameters in model 5. Shaded range indicates inter-quartile  interval of the population}
    \label{fig:model5_para}

\end{figure}

\begin{figure}[h]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/inferno5.png}}
    \end{center}
    \caption[Joint density distribution of each parameter pair]%
    {Joint distribution of each parameter pair. The diagonal is KDE fit of each individual parameter, the upper triangle contains scatter plot of the particles and the lower triangle is the joint density distribution for each parameter pair}
    \label{fig:model5_para_joint}
\end{figure}

\begin{table}[H]
    \centering
    \begin{tabular}{|c c c|}
        \hline
        Parameter            & Mean value & Inter-quartile interval \\[0.5ex]
        \hline\hline
        $\lambda_N$          & 13.5       & [12.9, 14.2]            \\
        $a$                  & 0.0175     & [0.0145, 0.0207]        \\
        $\kappa_{N\beta}$    & 0.176      & [0.109, 0.234]          \\
        $\mu_N$              & 0.0239     & [0.0113, 0.0332]        \\
        $\nu_{N\Phi}$        & 0.125      & [0.115, 0.134]          \\
        \hline
        $\kappa_{\Phi\beta}$ & 0.123      & [0.115, 0.131]          \\
        $\mu_\Phi$           & 0.0234     & [0.0199, 0.0267]        \\
        \hline
        $s_{\beta N}$        & 2.38       & [2.02, 2.73]            \\
        $\mu_\beta$          & 2.67       & [2.24, 3.10]            \\
        \hline
        $s_{\alpha\Phi}$     & 0.574      & [0.295, 0.786]          \\
        $\mu_\alpha$         & 8.91       & [6.07, 11.1]            \\
        $f_{\beta\alpha}$    & 0.577      & [0.361, 0.738]          \\
        \hline
    \end{tabular}
    \caption[Inferred parameter values of model 5]
    {Inferred parameter values of model 5}
    \label{table:estimated5}
\end{table}

\subsection{Sensitivity of parameters}

% [why last PCs]

% [which parameters are most sensitive]

% [compared to credible intervals plot]

% [what does that mean]

A quick sensitivity analysis of the dynamical systems models can be quantified using principal component analysis (PCA) \cite{Toni}, where the first PCs corresponds to sloppy parameters while the last PCs corresponds to stiff parameters \cite{sensitivity}. PCA can only provide rough approximated sensitivity behaviour. By identifying the component loadings, the PCs can be visualised in pie charts and used to indicate the sensitivity.

\begin{figure}[H]
    \begin{center}
        \resizebox{1.0\hsize}{!}{\includegraphics{fig/pc.png}}
    \end{center}

    \caption[Explained total variance in original parameters by each PC]{Explained total variance in original parameters by each PC. From PC 5 on. little variance is explained by each PC, which is enlarged in Right}
    \label{fig:pc}
\end{figure}

As models 5 won in the model selection, further interest was to explore the parameter-to-model sensitivity of the resultant model. PCA was performed on the last generation of the inferred model 5, Figure \ref{fig:pc} shows the total explained variance by each PC. The first PC explained a major part of the variance (91.5\%), and the rest PCs explained less than 5\% of the variance each. PC 5 to PC 12 explained much less variance (less than 0.1\% of the total variance) than the first four PCs. Thus, they were considered to extend much narrower regions on the posterior distributions and could be used to identify the parameters that the model is most sensitive to.

\begin{figure}[ht!]
    \begin{center}
        \resizebox{0.7\hsize}{!}{\includegraphics{fig/pc_pie.png}}
    \end{center}

    \caption[Fraction of the length of PCs explained by parameters]{Fraction of the length of PCs explained by parameters. Portions less than 2\% are not labelled}
    \label{fig:pc_pie}
\end{figure}

Figure \ref{fig:pc_pie} shows the fractions of the PCs explained by individual parameters, and a sum of the fractions in the last 8 PCs, i.e. PC 5 to PC 12. The last PC (PC12) is mostly the linear combination of $a$ (from the equation of neutrophil) and $\mu_\Phi$ (from the equation of macrophage), to which the model is most sensitive. If we conclude the last 8 PCs, it can be found that the biggest portions ($v_{N\Phi}$, $a$, $\kappa_{N\beta}$, $\kappa_{\Phi\beta}$, $\mu_N$, $\mu_\Phi$, $\nu_{N\Phi}$) mostly come from the equation of neutrophil and macrophage i.e. equations of the cells (the un-exploded wedges in Figure \ref{fig:pc_pie} a); if we look at the last 8 PCs individually, the similar pattern can be observed. This might suggest that the model is more sensitive to cells' dynamic rather than cytokines' dynamics.

The credible intervals (Figure \ref{fig:credible}) of each parameter across generations also agreed with the above conclusion from PCA. The stiff parameters, e.g. $a$ and $\mu_N$, extend narrow credible intervals, while sloppy parameters, e.g. $\mu_\alpha$ and $s_{\alpha\Phi}$, have wider credible intervals across the last few generations.

To conclude, model 5 is most sensitive to the parameters in the equations for cells' dynamics (i.e. first two equations in Eqn.\ref{eq:model5}), especially to $a$ and $\mu_\Phi$.

% then $v_{N\Phi}$, $\mu_N)$ (from the equation of neutrophil) and $k_{\Phi\beta}$ (from the equation of macrophage) contribute most portions together with $a$ and $\mu_\Phi$. As a conclusion, model 5 is sensitive to changes in the above parameters, which all come from the equation of neutrophil and macrophage i.e. equations of the cells; it may suggests that given the observed data, the inferred model is more sensitive to cells' dynamic rather than cytokines' dynamics.

% Also, some visualisations of the last population agreed with conclusions from PCA. Approximated posteriors of the last population plotted in FIGURE and credible ranges of parameters across populations in FIGURE shows that the 5 parameters identified by PCA is indeed `stiff' parameters.